HO unification from object language to meta language

### 2024-03-15 paper/main.tex
latex

We represent a logic program run for a goal g as
a list steps $p$ of length $\mathcal{N}$ each made of a unification problem
between terms $\mathcal{S}_{p_l}$ and $\mathcal{S}_{p_r}$
taken from the set of all terms $\mathcal{S}$.
If a rule is used
twice in a run we make two copies of the terms (the predicates)
occurring in the head or premises. So a run is the conjunction of all
unification problems, where $\rho_0$ is the empty substitution
%
\newcommand{\solve}{\ensuremath{\mathrm{check}}}
\newcommand{\step}[4]{\ensuremath{\mathrm{step}(#1,#2,#3) \mapsto #4}}
\newcommand{\stepD}[4]{\ensuremath{#3 #1_{#2_l} \Uo #3 #1_{#2_r} \mapsto #4}}
\newcommand{\runD}[2]{\ensuremath{\bigwedge_{p = 1 \ldots #2} \step{#1}{p}{\rho_{(p-1)}}{\rho_{p}}}}
\newcommand{\run}[2]{\ensuremath{\mathrm{run}(#1,#2)}}
$$
\step{\mathcal{S}}{p}{\rho}{\rho'} := \stepD{\mathcal{S}}{p}{\rho}{\rho'}
$$
$$
\run{\mathcal{S}}{\mathcal{N}} := \runD{\mathcal{S}}{\mathcal{N}}
$$
$$
\bigwedge_{p = 1 \ldots \mathcal{N}}
\underbrace{
  \rho_{(p-1)}\mathcal{S}_{p_l} \Uo \rho_{(p-1)}\mathcal{S}_{p_l} \mapsto \rho_p
}_\mathrm{step}
$$

We simulate this run as where $\sigma_0$ is the empty substitution
%
\begin{gather*}
\mathcal{T} \times \mathbb{M} \times \mathbb{L} = \{ (t_j,m_j,l_j) | s_j \in \mathcal{S}, \C{s_j}{t_j}{m_j}{l_j} \}\\
\bigwedge_{p = 1 \ldots \mathcal{N}}
 \underbrace{
  \sigma_{(p-1)}\mathcal{T}_{p_l} \Ue \sigma_{(p-1)}\mathcal{T}_{p_r} \mapsto \sigma_p' \land
  \solve~(\mathbb{L},\sigma_p')\mapsto \sigma_p
 }_\mathrm{step'}\\
\D{\sigma_{\mathcal{N}}}{\mathbb{L}}{\rho_{\mathcal{N}}}
\end{gather*}

\begin{proposition}[Simulation fidelity]\label{prop:fidelity}
  $\forall p, \mathrm{step}(p,\rho) \mapsto \rho' \Leftrightarrow
              \mathrm{step'}(p,\sigma) \mapsto \sigma'$
\end{proposition}

>>>>

We represent a logic program run for a goal g as
a list steps $p$ of length $\mathcal{N}$ each made of a unification problem
between terms $\mathcal{S}_{p_l}$ and $\mathcal{S}_{p_r}$
taken from the set of all terms $\mathcal{S}$.
If a rule is used
twice in a run we make two copies of the terms (the predicates)
occurring in the head or premises. So a run is the conjunction of all
unification problems, where $\rho_0$ is the empty substitution
%
\newcommand{\solve}{\ensuremath{\mathrm{check}}}
\newcommand{\fstep}{\ensuremath{\mathrm{fstep}}}
\newcommand{\hstep}{\ensuremath{\mathrm{hstep}}}
\newcommand{\frun}{\ensuremath{\mathrm{frun}}}
\newcommand{\hrun}{\ensuremath{\mathrm{hrun}}}
\newcommand{\stepF}[4]{\ensuremath{\fstep(#1,#2,#3) \mapsto #4}}
\newcommand{\stepH}[5]{\ensuremath{\hstep(#1,#2,#3,#4) \mapsto #5}}
\newcommand{\stepFD}[4]{\ensuremath{#3 #1_{#2_l} \Uo #3 #1_{#2_r} \mapsto #4}}
\newcommand{\stepHD}[6]{\ensuremath{%
#3 #1_{#2_l} \Ue #3 #1_{#2_r} \mapsto #4 \land \solve(#6,#4) \mapsto #5}}
\newcommand{\runF}[3]{\ensuremath{\frun(#1,#2) \mapsto #3_{#2}}}
\newcommand{\runH}[3]{\ensuremath{\hrun(#1,#2) \mapsto #3_{#2}}}
\newcommand{\runFD}[2]{\ensuremath{\bigwedge_{p = 1}^{#2} \stepF{#1}{p}{\rho_{p-1}}{\rho_{p}}}}
\newcommand{\runHD}[3]{\ensuremath{\bigwedge_{p = 1}^{#2} \stepH{#1}{p}{\sigma_{p-1}}{\sigma'}{\sigma_{p}}{#3}}}
\newcommand{\deff}{\ensuremath{\stackrel{{\scriptscriptstyle def}}{=\!=\!\!\!=}}}
$$
\begin{array}{l}
\stepF{\mathcal{S}}{p}{\rho}{\rho'}
\deff
\stepFD{\mathcal{S}}{p}{\rho}{\rho'}\\
\runF{\mathcal{S}}{\mathcal{N}}{\rho}
\deff
\runFD{\mathcal{S}}{\mathcal{N}}
\end{array}
$$

We simulate this run as where $\sigma_0$ is the empty substitution
$$
\begin{array}{l}
\stepH{\mathcal{T}}{p}{\sigma}{\mathbb{L}}{\sigma''} \deff\\
  \qquad\stepHD{\mathcal{T}}{p}{\sigma}{\sigma'}{\sigma''}{\mathbb{L}}\\
\runH{\mathcal{S}}{\mathcal{N}}{\rho} \deff \\
  \qquad \mathcal{T} \times \mathbb{M} \times \mathbb{L} = \{ (t_j,m_j,l_j) | s_j \in \mathcal{S}, \C{s_j}{t_j}{m_j}{l_j} \}\\
  \qquad \runFD{\mathcal{S}}{\mathcal{N}}{\mathbb{L}}\\
  \qquad \D{\sigma_{\mathcal{N}}}{\mathbb{M}}{\rho_{\mathcal{N}}}
\end{array}
$$

\begin{proposition}[Simulation fidelity]\label{prop:fidelity}
In the context of \hrun, we have that
$\forall p \in 1 \ldots \mathcal{N}$
$$
\stepF{\mathcal{S}}{p}{\rho_{p-1}}{\rho_p}
\Leftrightarrow
\stepH{\mathcal{T}}{p}{\sigma_{p-1}}{\mathbb{L}}{\sigma_p}
$$
\end{proposition}
